A tutorial of Bayesian beta regressions with brms in R

### 2025-02-05 manuscript/manuscript.tex
use tables in manuscript

Once we have the posterior difference, we can obtain CrIs of the
difference using \texttt{quantile2()} from the posterior package. Beware
that the values of the difference are in log-odds! We can convert these
into odd-ratios with \texttt{exp()}. odd-ratios indicate the ratio of
the difference between A and B, so that 1 means no difference, values
greater than 1 indicate an increase in A relative to B and values lower
than 1 indicate a decrease in A relative to B. odd-ratios are useful
when looking at differences that are in log-odds because while the
relative magnitude of the difference in proportion between two groups is
the same independent of the baseline proportion, the \emph{absolute}
magnitude of the difference depends on the baseline value. For example,
an odd-ratio difference of 1.25 would correspond to a proportion
increase of 13 percentage points if the baseline proportion is 0.62 but
it would correspond to a proportion increase of 17 percentage points if
the baseline proportion is 0.73. Of course, in real research contexts it
is still usefull to think about absolute magnitudes and their relevance
from a conceptual and methodological perspective. In this tutorial we
just focus on odd-ratios for simplicity.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(posterior)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
This is posterior version 1.6.0
\end{verbatim}

\begin{verbatim}

Attaching package: 'posterior'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:stats':

    mad, sd, var
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    %in%, match
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# round to 2 digits}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 2 x 3
    q90   q80   q60
  <dbl> <dbl> <dbl>
1  1.07  1.1   1.14
2  1.38  1.34  1.3 
\end{verbatim}

So, based on the model and data, there is a 90\% probability that the
voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
than in /t, k/. At 80\% confidence, the change ratio is 1.10-1.34 (or
10-34\% increase) while at 60\% confidence is 1.14-1.30 (14-30\%
increase). In other words we can be quite confident that the voicing
proportion in /p/ is longer than in /t, k/ and that the increase is less
than 35\%.

brms comes with a convenient function, \texttt{conditional\_effects()},
to plot posterior means and CrIs based on predictors in the model. In
the following example, we plot the predicted proportion of voicing by
consonant and vowel.

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}}

@@

\FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

          Contrast Estimate   10.0 % 90.0 %
 mean(t) - mean(k)   0.0348 -0.00418 0.0738
 mean(p) - mean(k)   0.2176  0.17837 0.2558
 mean(p) - mean(t)   0.1829  0.14474 0.2192

Term: c2
Type:  link 
Columns: term, contrast, estimate, conf.low, conf.high, predicted_lo, predicted_hi, predicted, tmp_idx 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

 vowel Estimate 10.0 % 90.0 %
     i    0.331  0.325  0.338
     e    0.300  0.293  0.307
     a    0.245  0.238  0.252
     o    0.305  0.298  0.312
     u    0.348  0.341  0.354

Type:  response 
Columns: vowel, estimate, conf.low, conf.high 
\end{verbatim}

@@

First, let's read and plot the data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
\NormalTok{nasal}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 705 x 6
   speaker label               vowel NC    voicing   nas_prop
   <chr>   <chr>               <chr> <chr> <chr>        <dbl>
 1 S03     b__U_nt_@___N_B17/s u     nt    voiceless  0.367  
 2 S03     b__a_nd_@___N_B19/s a     nd    voiced     0.195  
 3 S03     b__a_nt_@___N_B15/s a     nt    voiceless  0.279  
 4 S03     f__I_nt_@___N_B05/s i     nt    voiceless  0.764  
 5 S03     l__I_nd_@___N_B06/s i     nd    voiced     0.00529
 6 S03     p__E_nt_____N_B09/s e     nt    voiceless  0.335  
 7 S03     r__a_nt_@___N_B06/s a     nt    voiceless  0.243  
 8 S03     v__I_nd_@___N_B07/s i     nd    voiced     0.0248 
 9 S03     v__I_nt_6___N_B15/s i     nt    voiceless  0.135  
10 S03     z__E_nd_@___N_B17/s e     nd    voiced     0.538  
# i 695 more rows
\end{verbatim}

@@

@@

Let's inspect the output of \texttt{fixef()}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{round}\NormalTok{(}
  \FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
  \AttributeTok{digits =} \DecValTok{2}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
                        Estimate Est.Error   Q10   Q90
voicingvoiced:vowela       -0.61      0.12 -0.77 -0.46
voicingvoiceless:vowela    -0.08      0.09 -0.19  0.04
voicingvoiced:vowele       -0.19      0.16 -0.38  0.01
voicingvoiceless:vowele    -0.25      0.10 -0.38 -0.12
voicingvoiced:voweli       -0.74      0.19 -0.98 -0.51
voicingvoiceless:voweli    -0.12      0.18 -0.35  0.10
voicingvoiced:vowelo       -0.34      0.16 -0.54 -0.14
voicingvoiceless:vowelo     0.25      0.15  0.05  0.44
voicingvoiced:vowelu       -0.71      0.18 -0.94 -0.49
voicingvoiceless:vowelu    -0.58      0.16 -0.78 -0.38
\end{verbatim}

Negative log-odds indicate a proportion that is smaller than 50\%, while
positive log-odds a proportion that is greater than 50\%. Generally, the
expected log-odd predictions are negative, indicating an overall
tendency for the nasalisation to take less than 50\% of the duration of
the vowel. Moreover, the predictions are higher for voiceless NC
sequences than for voiced NC sequences, indicating a greater proportion
of nasalisation in the former. However there is vowel-specific
variation, and there doesn't seem to be much of a difference in
nasalisation proportion in /e/ and /u/. Plotting the expected
predictions with \texttt{conditional\_effects()} should make this
clearer.

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-1.pdf}}

@@

To answer question 1, we can calculate the average difference in
nasalisation proportion by first calculating the average nasalisation
across all vowels for voiced and voiceless sequences and then take the
difference of those, similarly to what we have done in the Case Study 1
above.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 40,000 x 7
   .chain .iteration .draw type  voicing          vowel    value
    <int>      <int> <int> <chr> <chr>            <chr>    <dbl>
 1      1          1     1 b     voicingvoiced    vowela -0.463 
 2      1          1     1 b     voicingvoiceless vowela -0.0944
 3      1          1     1 b     voicingvoiced    vowele -0.155 
 4      1          1     1 b     voicingvoiceless vowele -0.208 
 5      1          1     1 b     voicingvoiced    voweli -0.481 
 6      1          1     1 b     voicingvoiceless voweli -0.0529
 7      1          1     1 b     voicingvoiced    vowelo -0.310 
 8      1          1     1 b     voicingvoiceless vowelo  0.321 
 9      1          1     1 b     voicingvoiced    vowelu -0.620 
10      1          1     1 b     voicingvoiceless vowelu -0.368 
# i 39,990 more rows
\end{verbatim}

Now let's calculate the mean nasalisation proportion within each draw by
voicing, and plot the resulting posterior distributions. Note that, as
discussed for Case study 1, when working wit log-odds it is important to
first do all necessary calculations in log-odds, here calculate the mean
log-odds across vowels, and \emph{then} transform the calculated
estimands to proportions/probabilities.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \CommentTok{\# calculate the mean value within draw/voicing in log{-}odds}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value),}
    \CommentTok{\# we can now transform log{-}odds to proportion with plogis()}
    \AttributeTok{value\_mean\_prop =} \FunctionTok{plogis}\NormalTok{(value\_mean),}
    \AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_draws\_long\_voicing}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 8,000 x 4
   .draw voicing          value_mean value_mean_prop
   <int> <chr>                 <dbl>           <dbl>
 1     1 voicingvoiced       -0.406            0.400
 2     1 voicingvoiceless    -0.0804           0.480
 3     2 voicingvoiced       -0.391            0.404
 4     2 voicingvoiceless    -0.123            0.469
 5     3 voicingvoiced       -0.444            0.391
 6     3 voicingvoiceless    -0.132            0.467
 7     4 voicingvoiced       -0.439            0.392
 8     4 voicingvoiceless    -0.0982           0.475
 9     5 voicingvoiced       -0.383            0.405
10     5 voicingvoiceless    -0.266            0.434
# i 7,990 more rows
\end{verbatim}

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-plot-1.pdf}}

The plot suggests an overall greater nasalisation proportion in
voiceless NC sequences. Let's quantify how greater as we did in Case
Study 1. We will use odd-ratios in this context as well, i.e.~we will
convert log-odds to odd-ratios using the \texttt{exp()} (exponential)
function (and as before we first calculate the difference and then
exponentiate the resulting values, after which we can take summary
measures, like means and quantile-based measures such as CrIs).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
  \CommentTok{\# with one draw per raw. we need to drop the value\_mean\_prop col}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{value\_mean\_prop) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the difference of voiceless and voiced in log{-}odds}
    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced,}
    \CommentTok{\# now transform with exp() to get the ratio difference}
    \AttributeTok{voicing\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(voicing\_diff)}
\NormalTok{  )}
\NormalTok{nas\_prop\_bm\_diff}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 4,000 x 5
   .draw voicingvoiced voicingvoiceless voicing_diff voicing_diff_ratio
   <int>         <dbl>            <dbl>        <dbl>              <dbl>
 1     1        -0.406          -0.0804        0.325               1.38
 2     2        -0.391          -0.123         0.268               1.31
 3     3        -0.444          -0.132         0.312               1.37
 4     4        -0.439          -0.0982        0.340               1.41
 5     5        -0.383          -0.266         0.117               1.12
 6     6        -0.541          -0.0903        0.450               1.57
 7     7        -0.512          -0.111         0.401               1.49
 8     8        -0.493          -0.152         0.341               1.41
 9     9        -0.527          -0.175         0.352               1.42
10    10        -0.491          -0.257         0.234               1.26
# i 3,990 more rows
\end{verbatim}

>>>>

Once we have the posterior difference, we can obtain CrIs of the
difference using \texttt{quantile2()} from the posterior package. Beware
that the values of the difference are in log-odds! We can convert these
into odd-ratios with \texttt{exp()}. The output of the code below is
shown in Table~\ref{tbl-voi-bm-quant}. Odd-ratios indicate the ratio of
the difference between A and B, so that 1 means no difference, values
greater than 1 indicate an increase in A relative to B and values lower
than 1 indicate a decrease in A relative to B. odd-ratios are useful
when looking at differences that are in log-odds because while the
relative magnitude of the difference in proportion between two groups is
the same independent of the baseline proportion, the \emph{absolute}
magnitude of the difference depends on the baseline value. For example,
an odd-ratio difference of 1.25 would correspond to a proportion
increase of 13 percentage points if the baseline proportion is 0.62 but
it would correspond to a proportion increase of 17 percentage points if
the baseline proportion is 0.73. Of course, in real research contexts it
is still usefull to think about absolute magnitudes and their relevance
from a conceptual and methodological perspective. In this tutorial we
just focus on odd-ratios for simplicity.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(posterior)}

\NormalTok{voi\_bm\_quant }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# round to 2 digits}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-voi-bm-quant}Upper and lower limits of 90, 80 and
60\% Credible Intervals of the difference ratio of voicing proportion in
/t, k/ vs /p/.}

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.07 & 1.10 & 1.14 \\
1.38 & 1.34 & 1.30 \\

Based on the model and data, there is a 90\% probability that the
voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
than in /t, k/. At 80\% confidence, the change ratio is 1.10-1.34 (or
10-34\% increase) while at 60\% confidence is 1.14-1.30 (14-30\%
increase). In other words we can be quite confident that the voicing
proportion in /p/ is longer than in /t, k/ and that the increase is less
than 35\%. The brms package comes with a convenient function,
\texttt{conditional\_effects()}, to plot posterior means and CrIs based
on predictors in the model. In Figure~\ref{fig-voi-prop-bm-cond}, we
plot the predicted proportion of voicing by consonant and vowel.

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-cond-1.pdf}}

\caption{\label{fig-voi-prop-bm-cond}Expected voicing proportion by
consonant and vowel with 95\% Credible Intervals.}

@@

\FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
\FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}llrrr@{}}

\caption{\label{tbl-marginal-1}Difference in expected voicing proportion
by consonant, averaged across vowels, with 80\% CrIs.}

\toprule\noalign{}
term & contrast & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
c2 & mean(t) - mean(k) & 0.035 & -0.004 & 0.074 \\
c2 & mean(p) - mean(k) & 0.218 & 0.178 & 0.256 \\
c2 & mean(p) - mean(t) & 0.183 & 0.145 & 0.219 \\

\begin{longtable}[]{@{}lrrr@{}}

\caption{\label{tbl-marginal-2}Expected voicing proportion by vowel,
averaged across vowels, with 80\% CrIs.}

\toprule\noalign{}
vowel & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
i & 0.331 & 0.325 & 0.338 \\
e & 0.300 & 0.293 & 0.307 \\
a & 0.245 & 0.238 & 0.252 \\
o & 0.305 & 0.298 & 0.312 \\
u & 0.348 & 0.341 & 0.354 \\

@@

Table~\ref{tbl-nasal} shows the first ten rows of the data. The data
contains the following columns:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lllllr@{}}

\caption{\label{tbl-nasal}First ten rows of the nasal proportion data.}

\toprule\noalign{}
speaker & label & vowel & NC & voicing & nas\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
S03 & b\_\emph{U\_nt}@\_\_\_N\_B17/s & u & nt & voiceless & 0.3668820 \\
S03 & b\_\emph{a\_nd}@\_\_\_N\_B19/s & a & nd & voiced & 0.1954858 \\
S03 & b\_\emph{a\_nt}@\_\_\_N\_B15/s & a & nt & voiceless & 0.2786485 \\
S03 & f\_\emph{I\_nt}@\_\_\_N\_B05/s & i & nt & voiceless & 0.7642259 \\
S03 & l\_\emph{I\_nd}@\_\_\_N\_B06/s & i & nd & voiced & 0.0052949 \\
S03 & p\_\emph{E\_nt}\_\_\_\_N\_B09/s & e & nt & voiceless &
0.3347331 \\
S03 & r\_\emph{a\_nt}@\_\_\_N\_B06/s & a & nt & voiceless & 0.2431760 \\
S03 & v\_\emph{I\_nd}@\_\_\_N\_B07/s & i & nd & voiced & 0.0247572 \\
S03 & v\_\_I\_nt\_6\_\_\_N\_B15/s & i & nt & voiceless & 0.1350081 \\
S03 & z\_\emph{E\_nd}@\_\_\_N\_B17/s & e & nd & voiced & 0.5378522 \\

@@

Let's inspect the output of \texttt{fixef()}, reported in
Table~\ref{tbl-nas-prop-fixef}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\caption{\label{tbl-nas-prop-fixef}Regression coefficients of a beta
regression of nasalisation proportion (\texttt{nas\_prop\_bm}).}

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
voicingvoiced:vowela & -0.61 & 0.12 & -0.77 & -0.46 \\
voicingvoiceless:vowela & -0.08 & 0.09 & -0.19 & 0.04 \\
voicingvoiced:vowele & -0.19 & 0.16 & -0.38 & 0.01 \\
voicingvoiceless:vowele & -0.25 & 0.10 & -0.38 & -0.12 \\
voicingvoiced:voweli & -0.74 & 0.19 & -0.98 & -0.51 \\
voicingvoiceless:voweli & -0.12 & 0.18 & -0.35 & 0.10 \\
voicingvoiced:vowelo & -0.34 & 0.16 & -0.54 & -0.14 \\
voicingvoiceless:vowelo & 0.25 & 0.15 & 0.05 & 0.44 \\
voicingvoiced:vowelu & -0.71 & 0.18 & -0.94 & -0.49 \\
voicingvoiceless:vowelu & -0.58 & 0.16 & -0.78 & -0.38 \\

Negative log-odds indicate a proportion that is smaller than 50\%, while
positive log-odds a proportion that is greater than 50\%. Generally, the
expected log-odd predictions are negative, indicating an overall
tendency for the nasalisation to take less than 50\% of the duration of
the vowel. Moreover, the predictions are higher for voiceless NC
sequences than for voiced NC sequences, indicating a greater proportion
of nasalisation in the former. However there is vowel-specific
variation, and there doesn't seem to be much of a difference in
nasalisation proportion in /e/ and /u/. Figure~\ref{fig-nas-prop} shows
the expected predictions with \texttt{conditional\_effects()}, which
should make the interpretation clearer.

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-1.pdf}}

\caption{\label{fig-nas-prop}Expected nasalisation proportion by vowel
and voicing, with 95\% CrIs.}

@@

To answer question 1, we can calculate the average difference in
nasalisation proportion by first calculating the average nasalisation
across all vowels for voiced and voiceless sequences (see
Table~\ref{tbl-nas-draws} for the output of this step) and then take the
difference of those, similarly to what we have done in the Case Study 1
above.

@@

\caption{\label{tbl-nas-draws}First ten rows of posterior draws from
\texttt{nas\_prop\_bm} in long format.}

\toprule\noalign{}
.chain & .iteration & .draw & type & voicing & vowel & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voicingvoiced & vowela & -0.4632870 \\
1 & 1 & 1 & b & voicingvoiceless & vowela & -0.0943699 \\
1 & 1 & 1 & b & voicingvoiced & vowele & -0.1546258 \\
1 & 1 & 1 & b & voicingvoiceless & vowele & -0.2080810 \\
1 & 1 & 1 & b & voicingvoiced & voweli & -0.4810383 \\
1 & 1 & 1 & b & voicingvoiceless & voweli & -0.0529377 \\
1 & 1 & 1 & b & voicingvoiced & vowelo & -0.3096219 \\
1 & 1 & 1 & b & voicingvoiceless & vowelo & 0.3212456 \\
1 & 1 & 1 & b & voicingvoiced & vowelu & -0.6198312 \\
1 & 1 & 1 & b & voicingvoiceless & vowelu & -0.3680900 \\

Now let's calculate the mean nasalisation proportion within each draw by
voicing, and plot the resulting posterior distributions. Note that, as
discussed for Case study 1, when working wit log-odds it is important to
first do all necessary calculations in log-odds, here calculate the mean
log-odds across vowels, and \emph{then} transform the calculated
estimands to proportions/probabilities. The output of the following code
is show in Table~\ref{tbl-nas-long-voi} and the density plot of the
calculated draws is in
Figure~\ref{fig-nas-prop-bm-draws-long-voicing-plot}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \CommentTok{\# calculate the mean value within draw/voicing in log{-}odds}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value),}
    \CommentTok{\# we can now transform log{-}odds to proportion with plogis()}
    \AttributeTok{value\_mean\_prop =} \FunctionTok{plogis}\NormalTok{(value\_mean),}
    \AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rlrr@{}}

\caption{\label{tbl-nas-long-voi}Mean nasalisation proportion by draw
and voicing, averaged across vowels (first 10 rows).}

\toprule\noalign{}
.draw & voicing & value\_mean & value\_mean\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & voicingvoiced & -0.4056808 & 0.3999482 \\
1 & voicingvoiceless & -0.0804466 & 0.4798992 \\
2 & voicingvoiced & -0.3905744 & 0.4035790 \\
2 & voicingvoiceless & -0.1227124 & 0.4693603 \\
3 & voicingvoiced & -0.4436082 & 0.3908816 \\
3 & voicingvoiceless & -0.1320190 & 0.4670431 \\
4 & voicingvoiced & -0.4385607 & 0.3920840 \\
4 & voicingvoiceless & -0.0981768 & 0.4754755 \\
5 & voicingvoiced & -0.3827134 & 0.4054726 \\
5 & voicingvoiceless & -0.2657250 & 0.4339569 \\

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-bm-draws-long-voicing-plot-1.pdf}}

\caption{\label{fig-nas-prop-bm-draws-long-voicing-plot}Expected
nasalisation proportion by voicing, averaged across vowels.}

The plot suggests an overall greater nasalisation proportion in
voiceless NC sequences. Let's quantify how greater as we did in Case
Study 1. We will use odd-ratios in this context as well, i.e.~we will
convert log-odds to odd-ratios using the \texttt{exp()} (exponential)
function (and as before we first calculate the difference and then
exponentiate the resulting values, after which we can take summary
measures, like means and quantile-based measures such as CrIs).
tbl-nas-prop-bm-diff-quant show the 90, 80 and 60\% CrIs of the
difference ratio of nasalisation proportion in voiceless vs voiced NC
sequences.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
  \CommentTok{\# with one draw per raw. we need to drop the value\_mean\_prop col}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{value\_mean\_prop) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the difference of voiceless and voiced in log{-}odds}
    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced,}
    \CommentTok{\# now transform with exp() to get the ratio difference}
    \AttributeTok{voicing\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(voicing\_diff)}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_diff\_quant }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-nas-prop-bm-diff-quant}Upper and lower limits of 90,
80 and 60\% Credible Intervals of the difference ratio of nasalisation
proportion in vowels followed by voiced vs voiceless consonants.}

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.23 & 1.27 & 1.33 \\
1.69 & 1.63 & 1.56 \\

The CrIs of the ratio difference in nasalisation proportion in voiceless
vs voiced NC sequences suggest an increase of nasalisation in the
voiceless NC sequences, with a 90\% probability that the increase is
between 23\% and 69\% of the proportion in voiced NC sequences.
